ForOpenCL: Examing Fortran as a base language for accelerated platforms: I forgot what we decided upon

### 2010-12-03 opencl/papers/foropencl.tex
updating

@@

\title{ForOpenCL: Examing Fortran as a base language for accelerated platforms:
I forgot what we decided upon}

%
% \author{}
%

@@

\begin{abstract}

\end{abstract}

@@

@@

@@

The static analysis and source-to-source transformations used are not magic
and require the programmer to use a subset of Fortran that employs
a data-parallel programming model in this section we describe this
language subset.

@@

@@

Array sizes must be multiples of the local kernel size,
get_local_size(0)*get_local_size(1).  This may be relaxed in the future.

@@

   V = V - iA*(V - v_rest)                  !! Fortran

   V[l] = V[l] - iA[lex]*(V[l] - v_rest);   // OpenCL equivalent

@@

   where (V > Vth)                        !! Fortran
      iA = 1
   elsewhere
      iA = 0
   end where

   iA[lex] = (V[l] > Vth[l]) ? 1 : 0;     // OpenCL equivalent

@@

One of the state variables in the shallow water code is H, effectively the
height of the water.  This variable has a halo (ghost-cell region) surrounding
the interior of the grid to handle boundary conditions.  When using
MPI, the halo regions contains values from adjacent processors that must be
updated with new values at each time step.  This is accomplished with the
transfer_halo function.  The shallow water code assumes a five point stencil
so the state variables are extended by 2 in each dimension (e.g., by one to the
left, right, up, and down).

halo = [1,1,1,1]
iH => region(H, halo, transfer=false)

@@

   barrier(CLK_LOCAL_MEM_FENCE)

>>>>

\documentclass{article}

\title{ForOpenCL: Examing Fortran as a base language for accelerated platforms:
I forgot what we decided upon}
\author{Awesome Fortran Programmers}

@@

@@

\begin{abstract}
I am abstract.
\end{abstract}

@@

This paper presents a compiler-level approach for targeting a single program to multiple fundamentally different low level execution and programming models while allowing the application programmer to adopt a single high level programming model.

@@

\subsection{Why Fortran?}

A likely question that one may pose is ``\emph{Why Fortran and not a more modern language like X?}''
The recent rise in interest in concurrency and parallelism at the language level driven by multicore CPUs and
manycore accelerators has driven a number of new language developments, both as novel languages and extensions
on existing ones.

For scientific users, these new languages and language extensions present a challenge: how do developers effectively use them while avoiding rewriting their code and potentially growing dependent on a transient technology that will vanish tomorrow?

Fortran is unique in that it has contained language features that are well suited to modern architectures for a number of years.
This should be unsurprising --- Fortran was a primary language used to target systems such as the vector supercomputers and massively parallel systems of the 1970s and 1980s.  These are the systems in which architectural features were developed that have led to single chip high performance architectures of interest today.  Given that these new systems have features very similar to their predecessors, it is clear that the language features within Fortran for them are still relevant.

@@

The static analysis and source-to-source transformations used are very basic
and simply require the programmer to use a subset of Fortran that employs
a data-parallel programming model.  In particular, it encourages use of language features that
were introduced and standardized in the Fortran 90 language specification.  In this section we describe the
set of Fortran 90 features that our analysis and transformation method are based on.

@@

%% cites: ZPL, CM5 stencil compilers

% put something here on:
%
% shifts
% array notation
%

@@

Array sizes must be multiples of the local kernel size,
{\tt get\_local\_size(0)*get\_local\_size(1)}.  This may be relaxed in the future.

@@

\begin{verbatim}
   V = V - iA*(V - v_rest)                  !! Fortran

   V[l] = V[l] - iA[lex]*(V[l] - v_rest);   // OpenCL equivalent
\end{verbatim}

@@

\begin{verbatim}
   where (V > Vth)                        !! Fortran
      iA = 1
   elsewhere
      iA = 0
   end where

   iA[lex] = (V[l] > Vth[l]) ? 1 : 0;     // OpenCL equivalent
\end{verbatim}

@@

One of the state variables in the shallow water code is H, effectively the
height of the water.  This variable has a halo (ghost-cell region) surrounding
the interior of the grid to handle boundary conditions.  When using
MPI, the halo regions contains values from adjacent processors that must be
updated with new values at each time step.  This is accomplished with the
{\tt transfer\_halo} function.  The shallow water code assumes a five point stencil
so the state variables are extended by 2 in each dimension (e.g., by one to the
left, right, up, and down).

\begin{verbatim}
halo = [1,1,1,1]
iH => region(H, halo, transfer=false)
\end{verbatim}

@@

\begin{verbatim}
   barrier(CLK_LOCAL_MEM_FENCE)
\end{verbatim}
